---
title: "<strong> Beachwatch 2022 Annual Report </strong> <img src=\"www/MCS_Logo_Full_TwoTone_Ink_Small_Screen_RGB.png\" style=\"float: right;width:25%;\"/> "
author: "Author: Greg Wannell"
date: "Report generated: `r format(Sys.time(), '%d-%m-%Y')`"
output:
    html_document
runtime: shiny
header-includes:
  - \usepackage{graphicx}
  - \usepackage{fancyhdr}
  - \usepackage{fancy}
  - \setlength\headheight{50pt}
  - \usepackage{fontspec}
  - \setmainfont{Poppins}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{colortbl}
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = TRUE)

library(tidyverse)
library(knitr)
library(kableExtra)
library(plotly)
library(ggplot2)
library(shiny)

load("data/BW_2022_data_analysed.RData")

```
<br>
This report uses year-round Beachwatch data from 2022. All charts are 
interactive and can be hovered over to bring up additional information. 
Charts can be downloaded as individual .PNG files by clicking on the 'Download 
plot as a png' icon at the top of the chart. The full report can be exported as
a PDF by pressing (⌘ Command on mac) or (Ctrl on Windows) + P. This opens a 
Print dialog window where you can choose the option 'Save as PDF'. 
Any question please contact Greg Wannell at greg.wannell@mcsuk.org. 

### **Choose a country: **

```{r echo=FALSE}
selectInput("name", "Select the country you're interested in",
            choices = unique(engagement_stats$stretch_country),
            selected = engagement_stats$stretch_country[6])


output$selected_site <- renderText({
  paste("Stretch Country: ", input$name)
})

```

<br>

## **`r textOutput("selected_site")`**

### **Summary**

Summary statistics across all surveys within the selected region. 

<div style="margin-bottom:-20px;">
```{r echo=FALSE}

# get headline stats
output$t1 <- function() {
   engagement_stats %>% 
    ungroup() %>% 
    filter(stretch_country == input$name) %>% 
    select(n_surveys, volunteers, length_surveyed, bags_collected, weight_collected_kg) %>% 
      kbl(
      bottomrule = "",
      toprule = "",
      midrule = "",
      align = "l",
      digits = 1,
      col.names = c("Surveys", "Volunteers", "Length Surveyed", "Bag Collected", 
                "Weight Collected (kg)")) %>% 
    kable_styling()
}

tableOutput("t1")

```
</div>
```{r}
output$t2 <-  function() {
  litter_stats %>% 
  ungroup() %>% 
  filter(stretch_country == input$name) %>% 
    select(total_count, median_count) %>% 
    kbl(
  bottomrule = "",
  toprule = "",
  midrule = "",
  align = "l",
  digits = 1,
  col.names = c("Total items", "Average Items/100m (median)")) %>% 
  kable_styling(latex_options = "hover", position = "left", full_width = FALSE)
}

tableOutput("t2")
```

<br>

### **Top 10 litter items** 

This looks at the most abundant litter items found in all surveys. The ranking is determined by their median count. If two items have the same median count, the item 
with the higher total number of items found will be ranked higher.

```{r}
require(plotly)
  
output$p1 <- renderPlotly({
  top_10_median %>%
  ungroup() %>% 
  filter(stretch_country == input$name) %>% 
  select(litter_item, median_count, percent_of_median, presence) %>% 
  plot_ly(
  x = ~median_count,
  y = ~reorder(litter_item, median_count),
  type = "bar",
  orientation = 'h',
  marker = list(color = "#232332")) %>% 
  layout(yaxis = list(title = "", zeroline = FALSE, showline = FALSE),
         xaxis = list(title = "", showgrid = FALSE))
  
})

plotlyOutput("p1")

```

```{r fig.align='left', out.width='100%'}

output$t3 <-  function() {
  top_10_median %>%
    ungroup() %>% 
  filter(stretch_country == input$name) %>% 
  select(litter_item, median_count, percent_of_median, presence) %>% 
    kbl(
      col.names = c("Litter Item",
                              "Average/100m",
                              "% of Median",
                              "% of Cleans")) %>%
    kable_styling("hover", full_width = TRUE,
                  position = "left", )
}

tableOutput("t3")
```

<br>

### **Top 10 most widespread items**

This examines litter items that are regularly identified in surveys, regardless of the quantity found. Values are reported as percentages. 

```{r}
require(plotly)
  
output$p5 <- renderPlotly({
  top_10_presence %>%
  ungroup() %>% 
  filter(stretch_country == input$name) %>% 
  select(litter_item, presence) %>% 
  plot_ly(
  x = ~presence,
  y = ~reorder(litter_item, presence),
  type = "bar",
  orientation = 'h',
  marker = list(color = "#4dc9d6")) %>% 
  layout(yaxis = list(title = "", zeroline = FALSE, showline = FALSE),
         xaxis = list(title = "", showgrid = FALSE))
  
})

plotlyOutput("p5")

```

### **Sources** 

This looks at  the main sources of litter found across all surveys. Source of 
beach litter is a proxy for where that litter has likely derived from. Currently, 
MCS separates litter items into 7 sources which can help highlight spatial 
differences in human activity and habits, and the physical characteristics of 
each beach stretch/reporting region. 

```{r sources}

output$p2 <- renderPlotly({
  source_stats %>%
    ungroup() %>% 
  filter(stretch_country == input$name) %>% 
  select(source, total_count, median_count, percent_of_median, presence) %>% 
  rename("Source" = "source", 
         "Items found" = "total_count", 
         "Items per 100m" = "median_count",
         "% of Median" = "percent_of_median",
         "% of Cleans" = "presence") %>%
  plot_ly(
    labels = ~Source, 
              values = ~`% of Median`,
              textposition = '', 
              textinfo = 'percent', 
              textfont = list(color = '#000000', size = 12),
              text = ~paste(`Items per 100m`, " per 100m"), 
              hoverinfo = 'label+text',
              marker = list(line = list(color = "#FFFFFF", width = 2)), 
                            colors=c("#00b9b0", "#4dc9d6", "#a3d1ce", 
                                     "#38af87", "#518c87","#3d9be2", "blue4")) %>%
  add_pie(hole = 0.55) %>% 
  layout(title = '', showlegend = TRUE, legend = list(font = list(size = 12)),
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
})

plotlyOutput("p2")

```

```{r sources_table}

output$t4 <-  function() {
  source_stats %>%
     ungroup() %>% 
  filter(stretch_country == input$name) %>% 
  select(source, total_count, median_count, percent_of_median, presence) %>% 
  rename("Source" = "source", 
         "Items found" = "total_count", 
         "Items per 100m" = "median_count",
         "% of Median" = "percent_of_median",
         "% of Cleans" = "presence") %>% 
    kbl() %>%
    kable_styling("hover", full_width = TRUE,
                  position = "left", )
}

tableOutput("t4")
 
```
<br>

### **Materials**

This looks at the material composition of litter found across all surveys. 
Currently, there are 11 categories which refer to the main construction material 
the litter is made of. 

```{r materials}

output$p3 <- renderPlotly({
  material_stats %>%
     ungroup() %>% 
    filter(stretch_country == input$name) %>% 
    select(material, total_count, median_count, percent_of_median, presence) %>% 
    rename("Material" = "material", 
         "Items found" = "total_count", 
         "Items per 100m" = "median_count",
         "% of Median" = "percent_of_median",
         "% of Cleans" = "presence") %>%
    plot_ly(
      labels = ~Material,
      values = ~`% of Median`,
      textposition = '',
      textinfo = 'percent',
      textfont = list(color = '#000000', size = 12),
      text = ~paste(`Items per 100m`, " per 100m"),
      hoverinfo = 'label+text',
      marker = list(line = list(color = "#FFFFFF", width = 2)),
      colors=c("#00b9b0", "#4dc9d6", "#a3d1ce", 
                                     "#38af87", "#518c87","#3d9be2", "blue4")) %>%
    add_pie(hole = 0.55) %>%
    layout(title = '', showlegend = TRUE, legend = list(font = list(size = 12)),
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
})

plotlyOutput("p3")
```

```{r material_table}

output$t5 <-  function() {
  material_stats %>%
      ungroup() %>%
  filter(stretch_country == input$name) %>% 
  select(material, total_count, median_count, percent_of_median, presence) %>% 
  rename("Material" = "material", 
         "Items found" = "total_count", 
         "Items per 100m" = "median_count",
         "% of Median" = "percent_of_median",
         "% of Cleans" = "presence") %>% 
    kbl() %>%
    kable_styling("hover", full_width = TRUE,
                  position = "left", )
}

tableOutput("t5")
 
```