---
title: "<img src=\"www/MCS_Logo_Full_TwoTone_Ink_Small_Screen_RGB.png\" style=\"float: right;width:25%;\"/> "
output:
    html_document
runtime: shiny
header-includes:
  - \usepackage{graphicx}
  - \usepackage{fancyhdr}
  - \usepackage{fancy}
  - \setlength\headheight{28pt}
  - \usepackage{fontspec}
  - \setmainfont{Poppins}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{colortbl}
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

library(tidyverse)
library(knitr)
library(kableExtra)
library(plotly)
library(ggplot2)

load("data/BW_NW_2022_data_analysed.RData")

```

<br>
<br>

# **`r textOutput("selected_site")`**
#### Author: Greg Wannell
#### Report generated: `r Sys.Date()`

<br>
<br>

This report uses year-round Beachwatch data from the years 2022 to
2022. All charts are interactive and can be hovered over
to bring up additional information. Charts can be downloaded as individual 
.PNG files by clicking on the 'Download plot as a png' icon at the top of the 
chart. The full report can be exported as a PDF by pressing (⌘ Command on 
mac) or (Ctrl on Windows) + P. This opens a Print dialog window where you can 
choose the option 'Save as PDF'. Any question please contact Greg Wannell at 
greg.wannell@mcsuk.org. 

<br>

### **Choose a stretch: **

```{r echo=FALSE}
selectInput("name", "Select the stretch you're interested in",
            choices = unique(bw$stretch_name),
            selected = bw$stretch_name[1])


output$selected_site <- renderText({
  paste("Stretch Name: ", input$name)
})

```
## **Summary**

Summary statistics across all surveys within the given location and time frame. 

<div style="margin-bottom:-20px;">
```{r echo=FALSE}

# get headline stats

require(knitr)
require(kableExtra)

output$t1 <- function() {
   engagement_stats %>% 
  filter(stretch_name == input$name) %>% 
    select(n_surveys, volunteers, length_surveyed, bags_collected, weight_collected_kg) %>% 
      kbl(
      bottomrule = "",
      toprule = "",
      midrule = "",
      align = "l",
      digits = 1,
      col.names = c("Surveys", "Volunteers", "Length Surveyed", "Bag Collected", 
                "Weight Collected (kg)")) %>% 
    kable_styling()
}

tableOutput("t1")

```
</div>
```{r echo=FALSE}
output$t2 <-  function() {
  litter_stats %>% 
  ungroup() %>% 
  filter(stretch_name == input$name) %>% 
    select(total_count,mean_count) %>% 
    kbl(
  bottomrule = "",
  toprule = "",
  midrule = "",
  align = "l",
  digits = 1,
  col.names = c("Total items", "Average Items/100m")) %>% 
  kable_styling(latex_options = "hover", position = "left", full_width = FALSE)
}

tableOutput("t2")
```

<br>

## **Top 10 litter items** 

This looks at the most abundant litter items found across all surveys within the
given location and time frame. This information is used to help monitor the 
dominance of litter items and identify where work is needed. 

```{r include=TRUE}
require(plotly)
  
output$p1 <- renderPlotly({
  top_10 %>%
  filter(stretch_name == input$name) %>% 
  ungroup() %>% 
  select(litter_item, mean_count, percent_of_total, presence) %>% 
  plot_ly(
  x = ~mean_count,
  y = ~reorder(litter_item, mean_count),
  type = "bar",
  orientation = 'h',
  marker = list(color = "#232332")) %>% 
  layout(yaxis = list(title = "", zeroline = FALSE, showline = FALSE),
         xaxis = list(title = "", showgrid = FALSE))
  
})

plotlyOutput("p1")

```

```{r include=TRUE, echo=FALSE, fig.align='left', out.width='100%'}

output$t3 <-  function() {
  top_10 %>%
  filter(stretch_name == input$name) %>% 
  ungroup() %>% 
  select(litter_item, mean_count, percent_of_total, presence) %>% 
    kbl(
      col.names = c("Litter Item",
                              "Average/100m",
                              "% of Total",
                              "% of Cleans")) %>%
    kable_styling("hover", full_width = TRUE,
                  position = "left", )
}

tableOutput("t3")
```

<br>

## **Sources** 

This looks at  the main sources of litter found across all surveys. Source of 
beach litter is a proxy for where that litter has likely derived from. Currently, 
MCS separates litter items into 7 sources which can help highlight spatial 
differences in human activity and habits, and the physical characteristics of 
each beach stretch/reporting region. 

```{r}

output$p2 <- renderPlotly({
  source_stats %>%
  filter(stretch_name == input$name) %>% 
  ungroup() %>% 
    select(source, total_count, mean_count, percent_of_total, presence) %>% 
  rename("Source" = "source", 
         "Items found" = "total_count", 
         "Items per 100m" = "mean_count",
         "% of Total" = "percent_of_total",
         "% of Cleans" = "presence") %>%
  plot_ly(
    labels = ~Source, 
              values = ~`% of Total`,
              textposition = '', 
              textinfo = 'percent', 
              textfont = list(color = '#000000', size = 12),
              text = ~paste(`Items per 100m`, " per 100m"), 
              hoverinfo = 'label+text',
              marker = list(line = list(color = "#FFFFFF", width = 2)), 
                            colors=c("#00b9b0", "#4dc9d6", "#a3d1ce", 
                                     "#38af87", "#518c87","#3d9be2", "blue4")) %>%
  add_pie(hole = 0.55) %>% 
  layout(title = '', showlegend = TRUE, legend = list(font = list(size = 12)),
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
})

plotlyOutput("p2")
```

```{r}

output$t4 <-  function() {
  source_stats %>%
  filter(stretch_name == input$name) %>% 
  ungroup() %>% 
  select(source, total_count, mean_count, percent_of_total, presence) %>% 
  rename("Source" = "source", 
         "Items found" = "total_count", 
         "Items per 100m" = "mean_count",
         "% of Total" = "percent_of_total",
         "% of Cleans" = "presence") %>% 
    kbl() %>%
    kable_styling("hover", full_width = TRUE,
                  position = "left", )
}

tableOutput("t4")
 
```
<br>

## **Materials**

This looks at the material composition of the litter found across all surveys. 
Currently, there are 11 categories which refer to the main construction material 
the litter is made of. 

```{r}

output$p3 <- renderPlotly({
  material_stats %>%
    filter(stretch_name == input$name) %>% 
    ungroup() %>% 
    select(material, total_count, mean_count, percent_of_total, presence) %>% 
    rename("Material" = "material", 
         "Items found" = "total_count", 
         "Items per 100m" = "mean_count",
         "% of Total" = "percent_of_total",
         "% of Cleans" = "presence") %>%
    plot_ly(
      labels = ~Material,
      values = ~`% of Total`,
      textposition = '',
      textinfo = 'percent',
      textfont = list(color = '#000000', size = 12),
      text = ~paste(`Items per 100m`, " per 100m"),
      hoverinfo = 'label+text',
      marker = list(line = list(color = "#FFFFFF", width = 2)),
      colors=c("#00b9b0", "#4dc9d6", "#a3d1ce", 
                                     "#38af87", "#518c87","#3d9be2", "blue4")) %>%
    add_pie(hole = 0.55) %>%
    layout(title = '', showlegend = TRUE, legend = list(font = list(size = 12)),
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
})

plotlyOutput("p3")
```

```{r}

output$t5 <-  function() {
  material_stats %>%
  filter(stretch_name == input$name) %>% 
  ungroup() %>% 
  select(material, total_count, mean_count, percent_of_total, presence) %>% 
  rename("Material" = "material", 
         "Items found" = "total_count", 
         "Items per 100m" = "mean_count",
         "% of Total" = "percent_of_total",
         "% of Cleans" = "presence") %>% 
    kbl() %>%
    kable_styling("hover", full_width = TRUE,
                  position = "left", )
}

tableOutput("t5")
 
```