---
title: "<strong> Beachwatch 2022 Annual Report </strong> <img src=\"www/MCS_Logo_Full_TwoTone_Ink_Small_Screen_RGB.png\" style=\"float: right;width:25%;\"/> "
author: "Author: Greg Wannell"
date: "Last updated: `r format(Sys.time(), '%d-%m-%Y')`"
output:
    html_document
runtime: shiny
header-includes:
  - \usepackage{graphicx}
  - \usepackage{fancyhdr}
  - \usepackage{fancy}
  - \setlength\headheight{50pt}
  - \usepackage{fontspec}
  - \setmainfont{Poppins}
  - \usepackage{float}
  - \usepackage{booktabs}
  - \usepackage{colortbl}
---

```{r setup, echo=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, error = TRUE)

library(tidyverse)
library(knitr)
library(kableExtra)
library(plotly)
library(ggplot2)
library(shiny)
library(leaflet)
library(leaflet.extras)
library(htmltools)
library(tidyselect)

load("data/BW_2022_data_analysed.RData")

```
<br>
This report uses year-round Beachwatch data from 2022. All charts are 
interactive and can be hovered over to bring up additional information. 
Charts can be downloaded as individual .PNG files by clicking on the 'Download 
plot as a png' icon at the top of the chart. The full report can be exported as
a PDF by pressing (⌘ Command on mac) or (Ctrl on Windows) + P. This opens a 
Print dialog window where you can choose the option 'Save as PDF'. 
Any question please contact Greg Wannell at greg.wannell@mcsuk.org. 

### **Choose a country: **

```{r echo=FALSE}
selectInput("name", "Select the country you're interested in",
            choices = unique(engagement_stats$stretch_country),
            selected = engagement_stats$stretch_country[6])


output$selected_site <- renderText({
  paste("Stretch Country: ", input$name)
})

```

<br>

## **`r textOutput("selected_site")`**

### **Summary**

Summary statistics across all surveys within the selected region. 

<div style="margin-bottom:-20px;">
```{r echo=FALSE}

# get headline stats
output$t1 <- function() {
   engagement_stats %>% 
    ungroup() %>% 
    filter(stretch_country == input$name) %>% 
    select(n_surveys, volunteers, length_surveyed, bags_collected, weight_collected_kg) %>% 
      kbl(
      bottomrule = "",
      toprule = "",
      midrule = "",
      align = "l",
      digits = 1,
      col.names = c("Surveys", "Volunteers", "Length Surveyed", "Bag Collected", 
                "Weight Collected (kg)")) %>% 
    kable_styling()
}

tableOutput("t1")

```
</div>
```{r}
output$t2 <-  function() {
  litter_stats %>% 
  ungroup() %>% 
  filter(stretch_country == input$name) %>% 
    select(total_count, median_count) %>% 
    kbl(
  bottomrule = "",
  toprule = "",
  midrule = "",
  align = "l",
  digits = 1,
  col.names = c("Total items", "Average Items/100m (median)")) %>% 
  kable_styling(latex_options = "hover", position = "left", full_width = FALSE)
}

tableOutput("t2")
```

<br>

### **Top 10 litter items** 

This looks at the most abundant litter items found in all surveys. The ranking is determined by their median count. If two items have the same median count, the item 
with the higher total number of items found will be ranked higher.

```{r}
require(plotly)
  
output$p1 <- renderPlotly({
  top_10_median %>%
  ungroup() %>% 
  filter(stretch_country == input$name) %>% 
  select(litter_item, median_count, percent_of_median, presence) %>% 
  plot_ly(
  x = ~median_count,
  y = ~reorder(litter_item, median_count),
  text = ~ median_count,
  textposition = "inside",
  type = "bar",
  orientation = 'h',
  marker = list(color = "#232332")) %>% 
  layout(yaxis = list(title = "", zeroline = FALSE, showline = FALSE),
         xaxis = list(title = "", showgrid = FALSE))
  
})

plotlyOutput("p1")

```

<br>

### **Top 10 most widespread items**

This examines litter items that are regularly identified in surveys, regardless of the quantity found. Values are reported as percentages. 

```{r}
require(plotly)
  
output$p5 <- renderPlotly({
  top_10_presence %>%
  ungroup() %>% 
  filter(stretch_country == input$name) %>% 
  select(litter_item, presence) %>% 
  plot_ly(
  x = ~presence,
  y = ~reorder(litter_item, presence),
  text = ~ paste0(round(presence, 0), "%"),
  textposition = "inside",
  type = "bar",
  orientation = 'h',
  marker = list(color = "#4dc9d6")) %>% 
  layout(yaxis = list(title = "", zeroline = FALSE, showline = FALSE),
         xaxis = list(title = "", showgrid = FALSE))
  
})

plotlyOutput("p5")

```

<br>

```{r fig.align='left', out.width='100%'}

output$t3 <-  function() {
  item_stats %>%
    ungroup() %>% 
  filter(stretch_country == input$name) %>% 
    arrange(desc(median_count, total_count)) %>%
  select(litter_item, median_count, percent_of_median, presence) %>% 
    kbl(
      col.names = c("Litter Item",
                              "Average/100m",
                              "% of Median",
                              "% of Cleans")) %>%
    kable_styling("hover", full_width = TRUE,
                  position = "left", ) %>% 
    scroll_box(height = "400px", width = "100%")
}

tableOutput("t3")
```

### **Litter Source** 

This looks at  the main sources of litter found across all surveys. Source of 
beach litter is a proxy for where that litter has likely derived from. Currently, 
MCS separates litter items into 7 sources which can help highlight spatial 
differences in human activity and habits, and the physical characteristics of 
each beach stretch/reporting region. 

```{r sources}

output$p2 <- renderPlotly({
  source_stats %>%
    ungroup() %>% 
  filter(stretch_country == input$name) %>% 
  select(source, total_count, median_count, percent_of_median, presence) %>% 
  rename("Source" = "source", 
         "Items found" = "total_count", 
         "Items per 100m" = "median_count",
         "% of Median" = "percent_of_median",
         "% of Cleans" = "presence") %>%
  plot_ly(
    labels = ~Source, 
              values = ~`% of Median`,
              textposition = '', 
              textinfo = 'percent', 
              textfont = list(color = '#000000', size = 12),
              text = ~paste(`Items per 100m`, " per 100m"), 
              hoverinfo = 'label+text',
              marker = list(line = list(color = "#FFFFFF", width = 2)), 
                            colors=c("#00b9b0", "#4dc9d6", "#a3d1ce", 
                                     "#38af87", "#518c87","#3d9be2", "blue4")) %>%
  add_pie(hole = 0.55) %>% 
  layout(title = '', showlegend = TRUE, legend = list(font = list(size = 12)),
         xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
         yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
})

plotlyOutput("p2")

```

```{r sources_table}

output$t4 <-  function() {
  source_stats %>%
     ungroup() %>% 
  filter(stretch_country == input$name) %>% 
  select(source, total_count, median_count, percent_of_median, presence) %>% 
  rename("Source" = "source", 
         "Items found" = "total_count", 
         "Items per 100m" = "median_count",
         "% of Median" = "percent_of_median",
         "% of Cleans" = "presence") %>% 
    kbl() %>%
    kable_styling("hover", full_width = TRUE,
                  position = "left", )
}

tableOutput("t4")
 
```
<br>

### **Material Composition**

This looks at the material composition of litter found across all surveys. 
Currently, there are 11 categories which refer to the main construction material 
the litter is made of. 

```{r materials}

output$p3 <- renderPlotly({
  material_stats %>%
     ungroup() %>% 
    filter(stretch_country == input$name) %>% 
    select(material, total_count, median_count, percent_of_median, presence) %>% 
    rename("Material" = "material", 
         "Items found" = "total_count", 
         "Items per 100m" = "median_count",
         "% of Median" = "percent_of_median",
         "% of Cleans" = "presence") %>%
    plot_ly(
      labels = ~Material,
      values = ~`% of Median`,
      textposition = '',
      textinfo = 'percent',
      textfont = list(color = '#000000', size = 12),
      text = ~paste(`Items per 100m`, " per 100m"),
      hoverinfo = 'label+text',
      marker = list(line = list(color = "#FFFFFF", width = 2)),
      colors=c("#00b9b0", "#4dc9d6", "#a3d1ce", 
                                     "#38af87", "#518c87","#3d9be2", "blue4")) %>%
    add_pie(hole = 0.55) %>%
    layout(title = '', showlegend = TRUE, legend = list(font = list(size = 12)),
           xaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE),
           yaxis = list(showgrid = FALSE, zeroline = FALSE, showticklabels = FALSE))
})

plotlyOutput("p3")
```

```{r material_table}

output$t5 <-  function() {
  material_stats %>%
      ungroup() %>%
  filter(stretch_country == input$name) %>% 
  select(material, total_count, median_count, percent_of_median, presence) %>% 
  rename("Material" = "material", 
         "Items found" = "total_count", 
         "Items per 100m" = "median_count",
         "% of Median" = "percent_of_median",
         "% of Cleans" = "presence") %>% 
    kbl() %>%
    kable_styling("hover", full_width = TRUE,
                  position = "left", )
}

tableOutput("t5")
 
```

<br>
<br>
<br>
<br>
<br>
<br>
<br>
<br>

### **Choropleth map**

Use these choropleth maps to explore the spatial prevalence of key items. 
This makes it easy to identify any hotspots around the UK coastline. The map is 
divided into county regions. 

```{r echo=FALSE, out.width= '100%', out.height= 1000, warning=FALSE}

# Polygon highlighting options
poly_highlighting = highlightOptions(
  weight = 3,
  color = "white",
  fillOpacity = 1,
  bringToFront = TRUE
)

# Tooltip (label) options
tooltip_options = labelOptions(
  direction = "auto",
  textsize = "15px",
  opacity = 0.9,
  style = list(
    "font-family" = "Arial",
    "padding" = "6px",
    "border-color" = "black"
  )
)

# Colour bins
bins_SRD = c(0, 1, 5, 10, 20, 50, 100, 350)
bins_DRS = c(0, 1, 5, 10, 20, 50, 100, 200)
bins_SUP = c(0, 1, 5, 10, 20, 50, 100, 350)
bins_wet_wipe = c(0, 1, 5, 10, 20, 50, 100, 350)

SRD_labels = c("0", "1-2", "2-5", "5-10", "10-20", "20-50", "50-100", "100-350")
DRS_labels = c("0", "1-2", "2-5", "5-10", "10-20", "20-50", "50-100", "100-350")
SUP_labels = c("0", "1-2", "2-5", "5-10", "10-20", "20-50", "50-100", "100-350")
wet_wipe_labels = c("0", "1-2", "2-5", "5-10", "10-20", "20-50", "50-100", "100-350")

# Define colour palettes and bins
pal_SRD = colorBin("YlOrRd", domain = key_stats_map$SRD_median_count, bins = bins_SRD)
pal_DRS = colorBin("YlOrRd", domain = key_stats_map$DRS_median_count, bins = bins_DRS)
pal_SUP = colorBin("YlOrRd", domain = key_stats_map$SUP_median_count, bins = bins_SUP)
pal_wet_wipe = colorBin("YlOrRd", domain = key_stats_map$wet_wipe_median_count, bins = bins_wet_wipe)

add_legend = function(map, var_id, var_pal, var_title, ...){
  addLegend(
    map = map,
    data = key_stats_map,
    opacity = 0.7,
    na.label = "No data",
    position = "bottomright",
    layerId = var_id,
    title = var_id,
    pal = var_pal,
    # Required in order to select the correct variable via var_id
    values = select(key_stats_map, starts_with(str_sub(var_id, 1, 10))),
    ...
  )
}

# Function to add variable choropleths to map
add_choropleth = function(map, var_id, var_pal, ...){
  # Variable ID
  var_colname = vars_select(colnames(key_stats_map), starts_with(str_sub(var_id, 1, 10)))
  # Tooltip (label)
  tooltip = sprintf(
    "<b> %s </b> <br> Surveys: %g <br> Median: %s /100m %s",
    key_stats_map$CTYUA22NM,
    key_stats_map$year,
    key_stats_map[[var_colname]],
    ifelse(var_colname == "SRD_median_count", "/100m",
           ifelse(var_colname == "DRS_median_count", "/100m",
                  ifelse(var_colname == "SUP_median_count", "/100m",
                         ifelse(var_colname == "wet_wipe_median_count", "/100m"))))
  ) %>% lapply(htmltools::HTML)
  # Polygons
  addPolygons(
    map = map,
    data = key_stats_map,
    fillOpacity = 1,
    color = "black",
    weight = 1,
    opacity = 0.4,
    highlightOptions = poly_highlighting,
    label = tooltip,
    labelOptions = tooltip_options,
    group = var_id,
    # Required in order to select the correct variable via var_id
    fillColor = ~var_pal(key_stats_map[[var_colname]]),
    ...
  )
}

# Variable IDs
SRD = "SRD_median (Average/100m)"
DRS = "DRS_median (Average/100m)"
SUP = "SUP_median (Average/100m)"
wet_wipe = "wet_wipe_median (Average/100m)"

# Legend IDs
SRD_title = "Sewage Related Debris (SRD)"
DRS_title = "Deposit Return Scheme (DRS)"
SUP_title = "Single Use Plastics (SUP)"
wet_wipe_title = "Wet wipes"

# Plot map
m1 <-
  leaflet() %>%
  # Set view and zoom level
  setView(lng = -2.56,
          lat = 58.50,
          zoom = 6) %>% 
  # Set boundaries of map - scrolling outside will bounce you back to UK
  setMaxBounds(lng1 = -11.4, lat1 = 45.5,
               lng2 = 8, lat2 = 61.3) %>% 
  # Reset map to default setting
  addResetMapButton() %>% 
  # Add a scalebar
  addScaleBar(
    position = "bottomright",
    options = scaleBarOptions(imperial = FALSE)
  ) %>%
  # Choropleth polygons
  add_choropleth(var_id = SRD, var_pal = pal_SRD) %>% 
  add_choropleth(var_id = DRS, var_pal = pal_DRS) %>% 
  add_choropleth(var_id = SUP, var_pal = pal_SUP) %>%
  add_choropleth(var_id = wet_wipe, var_pal = pal_wet_wipe) %>%
  # Legends
  add_legend(var_id = SRD, var_pal = pal_SRD, var_title = SRD_leg) %>% 
  add_legend(var_id = DRS, var_pal = pal_DRS, var_title = DRS_leg) %>%
  add_legend(var_id = SUP, var_pal = pal_SUP, var_title = SUP_leg) %>% 
  add_legend(var_id = wet_wipe, var_pal = pal_wet_wipe, var_title = wet_wipe_leg) %>% 
  
  # Add layers control
  addLayersControl(
    options = layersControlOptions(collapsed = FALSE),
    baseGroups = c(SRD, wet_wipe, DRS, SUP)
  ) %>% 
  # Base group title
  htmlwidgets::onRender(
    jsCode = "function() { $('.leaflet-control-layers-base').prepend('<label style=\"text-align:left\"><strong><font size=\"4\">Key Items</font></strong><br>Average Per 100m</label>');}"
  ) %>% 
  # Switch legends when a different base group is selected
  # Code from here: https://gist.github.com/noamross/98c2053d81085517e686407096ec0a69
  htmlwidgets::onRender("
    function(el, x) {
      var initialLegend = 'SRD_median (Average/100m)' // Set the initial legend to be displayed by layerId
      var myMap = this;
      for (var legend in myMap.controls._controlsById) {
        var el = myMap.controls.get(legend.toString())._container;
        if(legend.toString() === initialLegend) {
          el.style.display = 'block';
        } else {
          el.style.display = 'none';
        };
      };
    myMap.on('baselayerchange',
      function (layer) {
        for (var legend in myMap.controls._controlsById) {
          var el = myMap.controls.get(legend.toString())._container;
          if(legend.toString() === layer.name) {
            el.style.display = 'block';
          } else {
            el.style.display = 'none';
          };
        };
      });
    }") %>% 
  # Add information icon (model onclick)
  # Code from here: https://stackoverflow.com/questions/68995343/r-leaflet-adding-an-information-popup-using-easybutton
  addBootstrapDependency() 
  #addEasyButton(easyButton(
   # icon = "fa-info-circle", title = "Map Information",
    #onClick = JS("function(btn, map){ $('#infobox').modal('show'); }")
  #)) %>% 
  #htmlwidgets::appendContent(info_content)

m1

```

```{r}
  downloadHandler(
    filename = paste0("heatmap", Sys.Date(), ".png"),
    content = function(file) {
      mapview::mapshot(m1(), file = file,
              cliprect = "viewport",
              selfcontained = FALSE)
    } # end of content() function
  ) # end of downloadHandler() function
```
